---
title: "FFR Data Matrix Creation"
author: "Jonas"
date: "2024-06-14"
output: html_document
---

Processes FFR data acquired using the extended Rapid technique. The input was preprocessed in isolated chunks, and this script prepares the data for further analysis.

=============================================================================
SETUP AND CONFIGURATION
=============================================================================

```{r setup, include=FALSE}
# Clear workspace and set chunk options
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load_packages, include=FALSE}
# Load required libraries with error handling
required_packages <- c("tidyverse")

lapply(required_packages, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
})
```

```{r configuration}
# Define file paths and parameters
CONFIG <- list(
  input_dir           = "./input/",
  output_dir          = "./output/",
  participant_ids     = 1:21,
  file_patterns       = list(
    snr_values        = "9MinRapidERPData_Part{id}Values.csv",
    signal_noise      = "9MinRapidERPData_Part{id}SignalNoiseValues.csv"
  ),
  harmonic_labels     = c("F0", "H2", "H3", "H4", "H5", "H6", "H7", "RMS", "HarmonicRMS"),
  signal_noise_labels = c("Signal", "Noise"),
  outlier_threshold   = 2.5  # Standard deviations for outlier detection
)
```

=============================================================================
UTILITY FUNCTIONS
=============================================================================

```{r utility-functions}
#' Create file path with participant ID
#' @param pattern File pattern with {id} placeholder
#' @param participant_id Participant ID to substitute
#' @return Complete file path
create_file_path <- function(pattern, participant_id) {
  gsub("\\{id\\}", participant_id, pattern)
}

#' Load participant data safely
#' @param file_path Path to CSV file
#' @return Data frame or NULL if file doesn't exist
load_participant_data <- function(file_path) {
  if (file.exists(file_path)) {
    read_csv(file_path, col_names = FALSE, show_col_types = FALSE)
  } else {
    warning(paste("File not found:", file_path))
    NULL
  }
}

#' Detect outliers using standard deviation method
#' @param data Numeric vector
#' @param threshold Number of standard deviations for outlier cutoff
#' @return Logical vector indicating outliers
detect_outliers <- function(data, threshold = 3) {
  mean_val    <- mean(data, na.rm = TRUE)
  sd_val      <- sd(data, na.rm = TRUE)
  upper_bound <- mean_val + threshold * sd_val
  lower_bound <- mean_val - threshold * sd_val
  
  data < lower_bound | data > upper_bound
}
```

=============================================================================
DATA LOADING FUNCTIONS
=============================================================================

```{r data-loading-functions}
#' Load all participant data for a specific data type
#' @param data_type Either "snr_values" or "signal_noise"
#' @return List of data frames
load_all_participant_data <- function(data_type) {
  setwd(CONFIG$input_dir)
  
  participant_data <- map(CONFIG$participant_ids, function(participant_id) {
    file_pattern   <- CONFIG$file_patterns[[data_type]]
    file_path      <- create_file_path(file_pattern, participant_id)
    load_participant_data(file_path)
  })
  
  # Remove NULL entries (failed loads)
  participant_data[!map_lgl(participant_data, is.null)]
}
```

=============================================================================
DATA PROCESSING FUNCTIONS
=============================================================================

```{r data-processing-functions}
#' Process SNR data into long format matrix
#' @param snr_data_list List of SNR data frames
#' @return Processed data frame in long format
process_snr_data <- function(snr_data_list) {
  if (length(snr_data_list) == 0) {
    stop("No SNR data loaded")
  }
  
  # Extract dimensions from first participant
  n_participants <- length(snr_data_list)
  n_chunks       <- nrow(snr_data_list[[1]])
  n_harmonics    <- as.numeric(snr_data_list[[1]][1, 1]) + 2
  
  # Create combinations for processing
  processing_grid <- expand_grid(
    listener    = 1:n_participants,
    chunk       = 1:n_chunks,
    harmonic    = 1:n_harmonics
  )
  
  # Process data using vectorized operations
  snr_matrix <- processing_grid %>%
    pmap_dfr(function(listener, chunk, harmonic) {
      data_frame <- snr_data_list[[listener]]
      
      if (harmonic <= n_harmonics - 2) {
        data_value <- data_frame[chunk, harmonic + 1]
        p_value    <- data_frame[chunk, harmonic + 10] + 0.000001
      } else {
        data_value <- data_frame[chunk, harmonic + 1]
        p_value    <- NA
      }
      
      tibble(
        listener   = as.character(listener),
        chunk      = as.character(chunk),
        harmonic   = harmonic,
        data_value = as.numeric(data_value),
        p_value    = as.numeric(p_value)
      )
    })
  
  # Apply factor transformations and labeling
  snr_matrix_processed <- snr_matrix %>%
    mutate(
      harmonic = factor(harmonic, 
                       levels = 1:length(CONFIG$harmonic_labels),
                       labels = CONFIG$harmonic_labels),
      chunk    = factor(chunk, levels = as.character(1:n_chunks)),
      listener = factor(listener)
    )
  
  return(snr_matrix_processed)
}

#' Process Signal/Noise data into long format matrix
#' @param signal_data_list List of signal/noise data frames
#' @return Processed data frame in long format
process_signal_noise_data <- function(signal_data_list) {
  if (length(signal_data_list) == 0) {
    stop("No signal/noise data loaded")
  }
  
  # Extract dimensions
  n_participants <- length(signal_data_list)
  n_chunks       <- nrow(signal_data_list[[1]])
  n_harmonics    <- as.numeric(signal_data_list[[1]][1, 1])
  
  # Create processing grid including signal/noise dimension
  processing_grid <- expand_grid(
    listener    = 1:n_participants,
    chunk       = 1:n_chunks,
    harmonic    = 1:n_harmonics,
    signal_type = 1:2  # 1 = Signal, 2 = Noise
  )
  
  # Process data
  signal_matrix <- processing_grid %>%
    pmap_dfr(function(listener, chunk, harmonic, signal_type) {
      data_frame <- signal_data_list[[listener]]
      
      if (signal_type == 1) {
        data_value <- data_frame[chunk, harmonic + 1]
      } else {
        data_value <- data_frame[chunk, harmonic + 8]
      }
      
      tibble(
        listener     = as.character(listener),
        chunk        = as.character(chunk),
        signal_noise = signal_type,
        harmonic     = harmonic,
        data_value   = as.numeric(data_value)
      )
    })
  
  # Apply factor transformations
  signal_matrix_processed <- signal_matrix %>%
    mutate(
      harmonic = factor(harmonic, 
                       levels = 1:length(CONFIG$harmonic_labels[1:7]),
                       labels = CONFIG$harmonic_labels[1:7]),
      chunk        = factor(chunk, levels = as.character(1:n_chunks)),
      listener     = factor(listener),
      signal_noise = factor(signal_noise, 
                           levels = 1:2, 
                           labels = CONFIG$signal_noise_labels)
    )
  
  return(signal_matrix_processed)
}
```

=============================================================================
OUTLIER DETECTION FUNCTIONS
=============================================================================

```{r outlier-detection-functions}
#' Identify outliers in signal/noise data
#' @param signal_data Processed signal/noise data frame
#' @return Data frame containing only outlier observations
identify_signal_outliers <- function(signal_data) {
  # Get unique combinations of harmonics and signal types
  conditions <- expand_grid(
    harmonic = levels(signal_data$harmonic),  
    signal_type = CONFIG$signal_noise_labels
  )
  
  # Find outliers for each condition
  outliers_list <- conditions %>%
    pmap(function(harmonic, signal_type) {
      condition_data <- signal_data %>%
        filter(harmonic == !!harmonic, signal_noise == !!signal_type)
      
      if (nrow(condition_data) > 0) {
        outlier_flags <- detect_outliers(condition_data$data_value, 
                                       CONFIG$outlier_threshold)
        condition_data[outlier_flags, ]
      } else {
        tibble()
      }
    })
  
  # Combine all outliers
  bind_rows(outliers_list)
}

#' Identify outliers in SNR data
#' @param snr_data Processed SNR data frame
#' @return Data frame containing only outlier observations
identify_snr_outliers <- function(snr_data) {
  harmonics_to_check <- CONFIG$harmonic_labels
  
  outliers_list <- map(harmonics_to_check, function(harmonic_level) {
    condition_data <- snr_data %>%
      filter(harmonic == harmonic_level)
    
    if (nrow(condition_data) > 0) {
      outlier_flags <- detect_outliers(condition_data$data_value, 
                                     CONFIG$outlier_threshold)
      condition_data[outlier_flags, ]
    } else {
      tibble()
    }
  })
  
  bind_rows(outliers_list)
}
```

=============================================================================
MAIN DATA PROCESSING PIPELINE
=============================================================================

```{r load-and-process-data}
# Load SNR data for all participants
cat("Loading SNR data...\n")
snr_raw_data <- load_all_participant_data("snr_values")
cat(paste("Loaded data for", length(snr_raw_data), "participants\n"))

# Load Signal/Noise data for all participants  
cat("Loading Signal/Noise data...\n")
signal_raw_data <- load_all_participant_data("signal_noise")
cat(paste("Loaded data for", length(signal_raw_data), "participants\n"))

# Process SNR data
cat("Processing SNR data...\n")
snr_matrix_final <- process_snr_data(snr_raw_data)
cat(paste("Created SNR matrix with", nrow(snr_matrix_final), "observations\n"))

# Process Signal/Noise data
cat("Processing Signal/Noise data...\n")
signal_matrix_final <- process_signal_noise_data(signal_raw_data)
cat(paste("Created Signal matrix with", nrow(signal_matrix_final), "observations\n"))

# Clean up intermediate objects
rm(snr_raw_data, signal_raw_data)
```

=============================================================================
OUTLIER DETECTION AND REMOVAL
=============================================================================

```{r outlier-detection}
# Identify outliers in signal data
cat("Identifying outliers in Signal/Noise data...\n")
signal_outliers <- identify_signal_outliers(signal_matrix_final)
cat(paste("Found", nrow(signal_outliers), "outliers in Signal/Noise data\n"))

# Identify outliers in SNR data
cat("Identifying outliers in SNR data...\n")
snr_outliers <- identify_snr_outliers(snr_matrix_final)
cat(paste("Found", nrow(snr_outliers), "outliers in SNR data\n"))

# Create clean datasets by removing outliers
snr_matrix_clean <- anti_join(snr_matrix_final, snr_outliers, 
                             by = c("listener", "chunk", "harmonic", 
                                   "data_value", "p_value"))

signal_matrix_clean <- anti_join(signal_matrix_final, signal_outliers,
                                by = c("listener", "chunk", "signal_noise", 
                                      "harmonic", "data_value"))

cat(paste("Clean SNR dataset:", nrow(snr_matrix_clean), "observations\n"))
cat(paste("Clean Signal dataset:", nrow(signal_matrix_clean), "observations\n"))
```

=============================================================================
DATA EXPORT
=============================================================================

```{r save-data}
# Set output directory
setwd(CONFIG$output_dir)

# Save original matrices
write_csv(snr_matrix_final, "SNRChunksMatrix.csv")
write_csv(signal_matrix_final, "SignalChunksMatrix.csv")

# Save cleaned matrices (without outliers)
write_csv(snr_matrix_clean, "SNRChunksMatrixClean.csv")
write_csv(signal_matrix_clean, "SignalChunksMatrixClean.csv")

# Save outlier datasets for inspection
write_csv(snr_outliers, "SNROutliers.csv")
write_csv(signal_outliers, "SignalOutliers.csv")

cat("All datasets saved successfully!\n")
```

=============================================================================
SUMMARY
=============================================================================

```{r data-summary}
# Summary statistics for final datasets
cat("=== SNR Data Summary ===\n")
print(summary(snr_matrix_final))

cat("\n=== Signal/Noise Data Summary ===\n") 
print(summary(signal_matrix_final))

# Check data structure
cat("\n=== Data Structure Validation ===\n")
cat("SNR matrix dimensions:", dim(snr_matrix_final), "\n")
cat("Signal matrix dimensions:", dim(signal_matrix_final), "\n")
cat("Unique participants in SNR data:", length(unique(snr_matrix_final$listener)), "\n")
cat("Unique participants in Signal data:", length(unique(signal_matrix_final$listener)), "\n")
```

=============================================================================
SESSION INFORMATION
=============================================================================

```{r session_info}
# Document the R session for reproducibility
sessionInfo()
```