---
title: "Long Rapid FFR Analysis"
author: "Jonas"
date: "2024-09-10"
output: html_document
---

Analyses the behaviour of the Rapid FFR recording method under continuous stimulation over a 9-minute period, assessing potential response saturation.
Research Question 4: Does prolonged stimulation cause saturation in the Rapid FFR response?

=============================================================================
SETUP AND CONFIGURATION
=============================================================================

```{r setup, include=FALSE}
# Clear environment for reproducible analysis
rm(list = ls())

# Load required packages with error handling
required_packages <- c("tidyverse", "lme4")

lapply(required_packages, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
})
```

```{r constants}
# Define analysis parameters
HARMONICS_OF_INTEREST <- c('F0', 'H2', 'H3', 'H4', 'H5', 'H6', 'H7')
HARMONIC_MAPPING      <- c("F0" = 1, "H2" = 2, "H3" = 3, "H4" = 4, "H5" = 5, "H6" = 6, "H7" = 7)
PARTICION_LEVELS      <- c('1', '2', '3', '4', '5', '6')
EFR_THRESHOLD         <- 4  # Harmonics below this are EFR, above are TFS
MAX_ITERATIONS        <- 1e6

# File paths (consider moving to config file)
DATA_PATH            <- "./input/"
SNR_INPUT_FILE       <- paste0(DATA_PATH, "SNRChunksMatrixClean.csv")
SIGNAL_INPUT_FILE    <- paste0(DATA_PATH, "SignalChunksMatrixClean.csv")
```

=============================================================================
DATA LOADING AND PREPROCESSING FUNCTIONS
=============================================================================

```{r data_functions}
#' Load and preprocess harmonic data
#' @param file_path Path to the CSV file
#' @return Cleaned and formatted dataframe
load_harmonic_data <- function(file_path) {
  tryCatch({
    data <- read_csv(file_path, show_col_types = FALSE) %>%
      mutate(
        data_value = as.numeric(data_value),
        listener = as.factor(listener)
      )
    
    return(data)
  }, error = function(e) {
    stop(paste("Error loading data from", file_path, ":", e$message))
  })
}

#' Convert harmonic labels to numeric values for analysis
#' @param data Dataframe containing harmonic data
#' @return Dataframe with numeric harmonic values
convert_harmonics_to_numeric <- function(data) {
  data %>%
    filter(harmonic %in% HARMONICS_OF_INTEREST) %>%
    mutate(
      harmonic_numeric = recode(harmonic, !!!HARMONIC_MAPPING),
      chunk_numeric = as.numeric(chunk)
    ) %>%
    group_by(chunk_numeric, harmonic_numeric) %>%
    mutate(
      mean_snr = mean(data_value, na.rm = TRUE),
      median_snr = median(data_value, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(listener, harmonic_numeric, chunk_numeric, data_value, mean_snr, median_snr)
}

#' Separate data into EFR and TFS components
#' @param data Processed harmonic data
#' @return List containing EFR and TFS datasets
separate_efr_tfs <- function(data) {
  list(
    efr = filter(data, harmonic_numeric < EFR_THRESHOLD),
    tfs = filter(data, harmonic_numeric >= EFR_THRESHOLD)
  )
}
```

=============================================================================
STATISTICAL MODELLING FUNCTIONS
=============================================================================

```{r modeling_functions}
#' Create optimal mixed-effects model using backward selection
#' @param data Dataset for modeling
#' @param response_var Response variable name
#' @return Optimal lmer model
create_optimal_model <- function(data, response_var = "data_value") {
  # Full model with quadratic terms and interactions
  full_formula <- as.formula(paste(
    response_var, 
    "~ (harmonic_numeric + I(harmonic_numeric^2)) * (chunk_numeric + I(chunk_numeric^2)) + (1|listener)"
  ))
  
  # Fit full model
  full_model <- lmer(
    full_formula,
    data       = data,
    REML       = FALSE,
    na.action  = na.fail,
    control    = lmerControl(optCtrl = list(maxfun = MAX_ITERATIONS))
  )
  
  # Backward selection to find optimal model
  optimal_model <- optimize_model_backward(full_model, data, response_var)
  
  return(optimal_model)
}

#' Perform backward model selection
#' @param full_model Starting full model
#' @param data Dataset
#' @param response_var Response variable name
#' @return Optimized model
optimize_model_backward <- function(full_model, data, response_var) {
  current_model <- full_model
  
  # Remove interaction terms if not significant
  reduced_model <- update(current_model, . ~ . - (chunk_numeric + I(chunk_numeric^2)):(harmonic_numeric + I(harmonic_numeric^2)))
  
  if (anova(current_model, reduced_model)$`Pr(>Chisq)`[2] > 0.05) {
    current_model <- reduced_model
  }
  
  # Test quadratic particion term
  test_model <- update(current_model, . ~ . - I(chunk_numeric^2))
  if (anova(current_model, test_model)$`Pr(>Chisq)`[2] > 0.05) {
    current_model <- test_model
  }
  
  # Test linear particion term
  test_model <- update(current_model, . ~ . - chunk_numeric)
  if (anova(current_model, test_model)$`Pr(>Chisq)`[2] > 0.05) {
    current_model <- test_model
  }
  
  return(current_model)
}

#' Generate model predictions and create visualization
#' @param model Fitted lmer model
#' @param data Original dataset
#' @param y_label Y-axis label for plot
#' @return List containing predictions and plot
create_model_predictions <- function(model, data, y_label) {
  # Generate predictions
  predictions_data <- data %>%
    mutate(predicted_values = predict(model, newdata = ., re.form = NA))
  
  # Create visualization
  prediction_plot <- ggplot(predictions_data, aes(x = chunk_numeric)) +
    geom_line(aes(y = predicted_values), color = "blue", size = 1) +
    geom_point(aes(y = data_value), alpha = 0.3) +
    geom_point(aes(y = mean_snr), shape = 21, size = 4, fill = "white") +
    geom_point(aes(y = median_snr), shape = 4, size = 4) +
    scale_x_continuous(name = "Partition") +
    scale_y_continuous(name = y_label) +
    facet_wrap(~harmonic_numeric, labeller = labeller(harmonic_numeric = function(x) paste("Harmonic", x))) +
    theme_bw() +
    theme(
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "lightgray")
    )
  
  return(list(
    data = predictions_data,
    plot = prediction_plot
  ))
}
```

=============================================================================
MAIN ANALYSIS PIPELINE
=============================================================================

```{r main_analysis}
# Load and preprocess data
cat("Loading SNR data...\n")
snr_raw_data <- load_harmonic_data(SNR_INPUT_FILE)

cat("Loading Signal data...\n") 
signal_raw_data <- load_harmonic_data(SIGNAL_INPUT_FILE)

# Prepare datasets for analysis
snr_analysis_data    <- convert_harmonics_to_numeric(snr_raw_data)
signal_analysis_data <- signal_raw_data %>%
  filter(signal_noise == "Signal") %>%
  convert_harmonics_to_numeric()

# Separate into EFR and TFS components
snr_components    <- separate_efr_tfs(snr_analysis_data)
signal_components <- separate_efr_tfs(signal_analysis_data)

cat("Data preprocessing complete.\n")
```

=============================================================================
SNR ANALYSIS
=============================================================================

```{r snr_analysis}
cat("Analyzing SNR data...\n")

# EFR SNR Analysis
cat("  - Fitting EFR SNR model...\n")
efr_snr_model   <- create_optimal_model(snr_components$efr, "data_value")
efr_snr_results <- create_model_predictions(efr_snr_model, snr_components$efr, "SNR (dB)")

cat("EFR SNR Model Summary:n")
summary(efr_snr_model)

# TFS SNR Analysis  
cat("  - Fitting TFS SNR model...\n")
tfs_snr_model   <- create_optimal_model(snr_components$tfs, "data_value")
tfs_snr_results <- create_model_predictions(tfs_snr_model, snr_components$tfs, "SNR (dB)")

cat("TFS SNR Model Summary:\n")
summary(tfs_snr_model)

# Display plots
efr_snr_results$plot + ggtitle("EFR SNR Analysis")
tfs_snr_results$plot + ggtitle("TFS SNR Analysis")
```

=============================================================================
SIGNAL ANALYSIS
=============================================================================

```{r signal_analysis}
cat("Analyzing Signal data...\n")

# EFR Signal Analysis
cat("  - Fitting EFR Signal model...\n")
efr_signal_model   <- create_optimal_model(signal_components$efr, "data_value")
efr_signal_results <- create_model_predictions(efr_signal_model, signal_components$efr, "Signal level (μV)")

cat("EFR Signal Model Summary:\n")
summary(efr_signal_model)

# TFS Signal Analysis
cat("  - Fitting TFS Signal model...\n")
tfs_signal_model   <- create_optimal_model(signal_components$tfs, "data_value")
tfs_signal_results <- create_model_predictions(tfs_signal_model, signal_components$tfs, "Signal level (μV)")

cat("TFS Signal Model Summary:\n")
summary(tfs_signal_model)

# Display plots
efr_signal_results$plot + ggtitle("EFR Signal Analysis")
tfs_signal_results$plot + ggtitle("TFS Signal Analysis")
```

=============================================================================
RESULTS EXPORT
=============================================================================

```{r export_results}
#' Prepare final results for export
#' @param efr_results EFR analysis results
#' @param tfs_results TFS analysis results
#' @return Combined dataframe with proper factor labels
prepare_export_data <- function(efr_results, tfs_results) {
  # Reverse mapping for harmonic labels
  harmonic_labels        <- names(HARMONIC_MAPPING)
  names(harmonic_labels) <- HARMONIC_MAPPING
  
  combined_data <- bind_rows(efr_results$data, tfs_results$data) %>%
    mutate(
      harmonic = recode_factor(harmonic_numeric, !!!harmonic_labels),
      chunk    = chunk_numeric
    ) %>%
    select(-c(harmonic_numeric, chunk_numeric) ) # Remove numeric version
  
  return(combined_data)
}

# Prepare and export results
cat("Preparing results for export...\n")

snr_final_results    <- prepare_export_data(efr_snr_results, tfs_snr_results)
signal_final_results <- prepare_export_data(efr_signal_results, tfs_signal_results)

# Export results
write_csv(snr_final_results, paste0(DATA_PATH, "SNRChunksMatrixPred.csv"))
write_csv(signal_final_results, paste0(DATA_PATH, "SignalChunksMatrixPred.csv"))

cat("Analysis complete! Results exported to:", DATA_PATH, "\n")
```

=============================================================================
SESSION INFORMATION
=============================================================================

```{r session_info}
# Document the R session for reproducibility
sessionInfo()
```