---
title: "FFR Analysis: Conventional vs Rapid Stimulation"
author: "Jonas"
date: "2025-05-26"
output: html_document
---

Analyses and compares SNR values across harmonics for two experimental conditions (Rapid vs. Conventional FFR).
Research Question 1: Are SNR values significantly different between Rapid and Conventional stimulation protocols?

=============================================================================
SETUP AND CONFIGURATION
=============================================================================

```{r setup, include = FALSE}
# Clear workspace for clean start
rm(list = ls())

# Load required packages with error handling
required_packages <- c("tidyverse", "nlme", "afex", "ggpubr", "lme4")

lapply(required_packages, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
})
```

=============================================================================
DATA LOADING AND PREPROCESSING
=============================================================================

```{r data_loading}
# Define file paths (consider using here::here() for better path management)
DATA_PATH   <- "./input/"
SNR_FILE    <- "SNRIterationsMatrixClean.csv"
SIGNAL_FILE <- "SignalIterationsMatrixClean.csv"

# Load and preprocess SNR data
load_and_clean_data <- function(file_path) {
  data <- read_csv(file_path, show_col_types = FALSE) 
  return(data)
}

snr_data_raw    <- load_and_clean_data(file.path(DATA_PATH, SNR_FILE))
signal_data_raw <- load_and_clean_data(file.path(DATA_PATH, SIGNAL_FILE))

# Calculate time matching between Rapid and Conventional conditions
# This ensures fair comparison by matching recording durations
calculate_time_matching <- function(data) {
  
  # Get maximum recording time for Rapid condition
  rapid_max_time <- data %>%
    filter(listener  == 1, 
           condition == "Rapid", 
           run       == "Run1", 
           harmonic  == "F0") %>%
    pull(recording_time) %>%
    max(na.rm = TRUE)
  
  # Find closest matching time point in Conventional condition
  conv_matching_times <- data %>%
    filter(listener   == 1, 
           condition  == "Conv", 
           run        == "Run1", 
           harmonic   == "F0") %>%
    mutate(time_diff  = abs(recording_time - rapid_max_time)) %>%
    filter(time_diff  == min(time_diff, na.rm = TRUE))
  
  recording_time_match   <- conv_matching_times$recording_time[1]
  stimulation_time_match <- conv_matching_times$stimulation_time[1]
  
  return(list(
    rapid_max_time         = rapid_max_time,
    recording_time_match   = recording_time_match,
    stimulation_time_match = stimulation_time_match
  ))
}

time_matching <- calculate_time_matching(snr_data_raw)
```

=============================================================================
DATA VISUALISATION
=============================================================================

```{r data_visualisation}
# Create overview plot comparing conditions
create_overview_plot <- function(data, time_match, harmonics_subset = c('F0','H2','H3','H4','H5', "H6", "H7")) {
  
  plot_data <- data %>%
    filter(harmonic %in% harmonics_subset,
           run == "Run1",
           (condition == "Rapid" & iteration == 100) | 
           (condition == "Conv" & recording_time == time_match$recording_time_match))
  
  overview_plot <- ggplot(plot_data, aes(x = harmonic, y = data_value, fill = condition)) + 
    geom_boxplot(width = 0.5, outlier.shape = NA) +
    geom_jitter(color = "black", size = 0.7, alpha = 0.4, 
                position = position_jitterdodge(jitter.width = 0.01)) +
    scale_x_discrete(name = "Harmonic") +
    scale_y_continuous(name = "SNR (dB)") +
    ggtitle("SNR Comparison: Conventional vs Rapid Stimulation") +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      legend.position = "bottom"
    )
  
  # Generate summary statistics
  summary_stats <- plot_data %>%
    group_by(harmonic, condition) %>%
    summarise(
      mean_snr       = mean(data_value, na.rm = TRUE),
      median_snr     = median(data_value, na.rm = TRUE),
      sd_snr         = sd(data_value, na.rm = TRUE),
      n_observations = n(),
      .groups        = 'drop'
    )
  
  return(list(plot = overview_plot, summary = summary_stats))
}

overview_results <- create_overview_plot(snr_data_raw, time_matching)
overview_results$plot
overview_results$summary
```

=============================================================================
DATA PREPARATION FOR STATISTICAL ANALYSIS
=============================================================================

```{r data_preparation}
# Prepare data for statistical modeling
prepare_analysis_data <- function(data, time_match, iteration_value = 100, 
                                harmonics_subset = c('F0','H2','H3','H4','H5', "H6", "H7")) {
  
  # Define harmonic mapping for continuous variable
  harmonic_mapping <- setNames(1:7, c("F0", "H2", "H3", "H4", "H5", "H6", "H7"))
  
  analysis_data <- data %>%
    filter(
      harmonic %in% harmonics_subset,
      run == "Run1",
      (condition == "Rapid" & iteration == iteration_value) | 
      (condition == "Conv" & recording_time == time_match$recording_time_match)
    ) %>%
    mutate(harmonic_numeric = harmonic_mapping[harmonic]) %>%
    group_by(condition, run, harmonic_numeric) %>%
    mutate(
      mean_snr   = mean(data_value, na.rm = TRUE),
      median_snr = median(data_value, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(listener, run, condition, harmonic_numeric, data_value, mean_snr, median_snr)
  
  return(analysis_data)
}

# Prepare datasets for different analyses
snr_analysis_data_full    <- prepare_analysis_data(snr_data_raw, time_matching, iteration_value = 100)
snr_analysis_data_partial <- prepare_analysis_data(snr_data_raw, time_matching, iteration_value = 53)

# Prepare signal and noise data
prepare_signal_noise_data <- function(data, time_match, signal_type, iteration_value = 100,
                                    harmonics_subset = c('F0','H2','H3','H4','H5', "H6", "H7")) {
  
  harmonic_mapping <- setNames(1:7, c("F0", "H2", "H3", "H4", "H5", "H6", "H7"))
  
  filtered_data <- data %>%
    filter(
      harmonic %in% harmonics_subset,
      run          == "Run1",
      signal_noise == signal_type,
      (condition   == "Rapid" & iteration == iteration_value) | 
      (condition   == "Conv" & recording_time == time_match$recording_time_match)
    ) %>%
    mutate(harmonic_numeric = harmonic_mapping[harmonic]) %>%
    group_by(condition, run, harmonic_numeric) %>%
    mutate(
      mean_value   = mean(data_value, na.rm = TRUE),
      median_value = median(data_value, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(listener, run, condition, harmonic_numeric, data_value, mean_value, median_value)
  
  return(filtered_data)
}

signal_analysis_data <- prepare_signal_noise_data(signal_data_raw, time_matching, "Signal")
noise_analysis_data  <- prepare_signal_noise_data(signal_data_raw, time_matching, "Noise")
```

=============================================================================
STATISTICAL MODELING FOR THE HARMONIC SLOPE
=============================================================================

```{r harmonic_slope_stats}
# Fit multiple model types to determine best functional form
fit_harmonic_models <- function(dataModel, response_var = "data_value", 
                               predictor_var = "harmonic_numeric") {
  
  # Linear model
  linear_formula <- as.formula(paste(response_var, "~ condition *", predictor_var, "+ (1|listener)"))
  model_linear   <- mixed(linear_formula, data = dataModel, method = "S", expand_re = TRUE)
  lmer_linear    <- lmer(linear_formula, data = dataModel)
  
  # Quadratic model
  quad_formula    <- as.formula(paste(response_var, "~ condition * poly(", predictor_var, ", 2) + (1|listener)"))
  model_quadratic <- mixed(quad_formula, data = dataModel, method = "S", expand_re = TRUE)
  lmer_quadratic  <- lmer(quad_formula, data = dataModel)
  
  # Logarithmic model
  log_formula   <- as.formula(paste(response_var, "~ condition * log10(", predictor_var, ") + (1|listener)"))
  model_log     <- mixed(log_formula, data = dataModel, method = "S", expand_re = TRUE)
  lmer_log      <- lmer(log_formula, data = dataModel)
  
  return(list(
    linear      = list(mixed = model_linear, lmer = lmer_linear),
    quadratic   = list(mixed = model_quadratic, lmer = lmer_quadratic),
    logarithmic = list(mixed = model_log, lmer = lmer_log)
  ))
}

# Fit models for SNR data
snr_models <- fit_harmonic_models(snr_analysis_data_full)

# Display model summaries
cat("\n=== LINEAR MODEL RESULTS ===\n")
print(snr_models$linear$mixed)

cat("\n=== QUADRATIC MODEL RESULTS ===\n")
print(snr_models$quadratic$mixed)

cat("\n=== LOGARITHMIC MODEL RESULTS ===\n")
print(snr_models$logarithmic$mixed)
```

=============================================================================
MODEL COMPARISON AND VISUALISATION
=============================================================================

```{r harmonic_slope_viz}
# Generate predictions for model comparison
generate_model_predictions <- function(dataModel, models) {
  
  predictions <- dataModel %>%
    mutate(
      pred_linear      = predict(models$linear$lmer, newdata = ., re.form = NA),
      pred_quadratic   = predict(models$quadratic$lmer, newdata = ., re.form = NA),
      pred_logarithmic = predict(models$logarithmic$lmer, newdata = ., re.form = NA)
    )
  
  return(predictions)
}

snr_predictions <- generate_model_predictions(snr_analysis_data_full, snr_models)

# Create model comparison plots
create_model_comparison_plot <- function(dataModel, model_type) {
  
  pred_column <- paste0("pred_", model_type)
  
  plot <- ggplot(dataModel, aes(x = harmonic_numeric, color = condition)) +
    geom_line(aes(y = dataModel[[pred_column]]), size = 1.2) +
    geom_point(aes(y = data_value), alpha = 0.6, size = 1) +
    geom_point(aes(y = mean_snr), shape = 21, size = 4, fill = "white", stroke = 1.5) +
    geom_point(aes(y = median_snr), shape = 4, size = 3, stroke = 1.5) +
    scale_x_continuous(name = "Harmonic Number", breaks = 1:7, 
                      labels = c("F0", "H2", "H3", "H4", "H5", "H6", "H7")) +
    scale_y_continuous(name = "SNR (dB)") +
    ggtitle(paste("SNR Analysis:", str_to_title(model_type), "Model")) +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
      legend.position = "bottom"
    ) +
    labs(caption = "Circles = Mean, X = Median, Points = Individual observations")
  
  return(plot)
}

# Generate comparison plots
plot_linear      <- create_model_comparison_plot(snr_predictions, "linear")
plot_quadratic   <- create_model_comparison_plot(snr_predictions, "quadratic")
plot_logarithmic <- create_model_comparison_plot(snr_predictions, "logarithmic")

print(plot_linear)
print(plot_quadratic)
print(plot_logarithmic)
```

=============================================================================
SIMPLIFY THE MODEL
=============================================================================

```{r stats_model}
original_model <- lmer(data_value ~ condition * 
                   (harmonic_numeric + I(harmonic_numeric^2)) + 
                   (1|listener),
            data = snr_analysis_data_full, REML=FALSE, na.action=na.fail,
            control = lmerControl(optCtrl = list(maxfun = 1e6)))
summary(original_model)

# remove the interactions
original_model1 <- update(original_model, .~. -condition:(harmonic_numeric+I(harmonic_numeric^2)))

anova(original_model, original_model1)
summary(original_model1) # This will be the final model architecture
```

=============================================================================
FINAL MODEL SELECTION AND ANALYSIS
=============================================================================

```{r stats_model_viz}
# Based on visual inspection and model fit, select quadratic as final model, 
# and use reduced model from previous code-chunk as template
fit_final_model <- function(dataModel) {
  
  # Final quadratic model with explicit polynomial terms for interpretability
  final_formula <- data_value ~ condition + harmonic_numeric + I(harmonic_numeric^2) + (1|listener)
  
  final_model_mixed <- mixed(final_formula, data = dataModel, 
                             method = "S", expand_re = TRUE)
  
  final_model_lmer <- lmer(final_formula, data = dataModel, REML = FALSE, 
                          na.action = na.fail,
                          control = lmerControl(optCtrl = list(maxfun = 1e6)))
  
  return(list(mixed = final_model_mixed, lmer = final_model_lmer))
}

final_model <- fit_final_model(snr_analysis_data_full)

cat("\n=== FINAL MODEL RESULTS ===\n")
print(final_model$mixed)
print(summary(final_model$lmer))

# Generate final predictions and visualization
final_predictions <- snr_analysis_data_full %>%
  mutate(predicted_snr = predict(final_model$lmer, newdata = ., re.form = NA))

final_plot <- ggplot(final_predictions, aes(x = harmonic_numeric, color = condition)) +
  geom_line(aes(y = predicted_snr), size = 1.5, alpha = 0.8) +
  geom_point(aes(y = data_value), alpha = 0.5, size = 1.2) +
  geom_point(aes(y = mean_snr), shape = 21, size = 4, fill = "white", stroke = 1.5) +
  geom_point(aes(y = median_snr), shape = 4, size = 3, stroke = 1.5) +
  scale_x_continuous(name = "Harmonic Number", breaks = 1:7, 
                    labels = c("F0", "H2", "H3", "H4", "H5", "H6", "H7")) +
  scale_y_continuous(name = "SNR (dB)") +
  scale_color_manual(values = c("Conv" = "#E31A1C", "Rapid" = "#1F78B4")) +
  ggtitle("Final Model: Quadratic Fit of SNR across Harmonics") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  labs(caption = "Circles = Mean, X = Median, Points = Individual observations\nLines represent quadratic model predictions")

final_plot
```

=============================================================================
MODEL DIAGNOSTICS AND ADDITIONAL ANALYSES
=============================================================================

```{r stats_model_summary}
# Extract model coefficients for interpretation
extract_model_summary <- function(model) {
  model_summary <- summary(model$lmer)
  coefficients <- data.frame(
    Term = rownames(model_summary$coefficients),
    Estimate = model_summary$coefficients[, "Estimate"],
    SE = model_summary$coefficients[, "Std. Error"],
    t_value = model_summary$coefficients[, "t value"],
    p_value = model_summary$coefficients[, "Pr(>|t|)"],
    stringsAsFactors = FALSE
  )
  
  return(list(
    coefficients = coefficients,
    aic = AIC(model$lmer),
    bic = BIC(model$lmer)
  ))
}

final_model_summary <- extract_model_summary(final_model)

cat("\n=== MODEL SUMMARY STATISTICS ===\n")
print(final_model_summary$coefficients)
cat("AIC:", final_model_summary$aic, "\n")
cat("BIC:", final_model_summary$bic, "\n")
```

============================================================================
SIMPLIFY THE MODEL FOR 52% DATASET
============================================================================

```{r stats_model52%}
original_model_partial <- lmer(data_value ~ condition * 
                                 (harmonic_numeric + I(harmonic_numeric^2)) + 
                                 (1|listener),
                               data = snr_analysis_data_partial, REML=FALSE, na.action=na.fail,
                               control = lmerControl(optCtrl = list(maxfun = 1e6)))
summary(original_model_partial)

# remove the interactions
original_model1_partial <- update(original_model_partial, .~. -condition:(harmonic_numeric+I(harmonic_numeric^2)))

anova(original_model_partial, original_model1_partial)
summary(original_model1_partial) # This will be the final model architecture
```

=============================================================================
FINAL MODEL SELECTION AND ANALYSIS FOR 52% DATASET
=============================================================================

```{r stats_model_viz52%}
final_model_partial <- fit_final_model(snr_analysis_data_partial)

cat("\n=== FINAL MODEL RESULTS ===\n")
print(final_model_partial$mixed)
print(summary(final_model_partial$lmer))

# Generate final predictions and visualization
final_predictions_partial <- snr_analysis_data_partial %>%
  mutate(predicted_snr = predict(final_model_partial$lmer, newdata = ., re.form = NA))

final_plot_partial <- ggplot(final_predictions_partial, aes(x = harmonic_numeric, color = condition)) +
  geom_line(aes(y = predicted_snr), size = 1.5, alpha = 0.8) +
  geom_point(aes(y = data_value), alpha = 0.5, size = 1.2) +
  geom_point(aes(y = mean_snr), shape = 21, size = 4, fill = "white", stroke = 1.5) +
  geom_point(aes(y = median_snr), shape = 4, size = 3, stroke = 1.5) +
  scale_x_continuous(name = "Harmonic Number", breaks = 1:7, 
                    labels = c("F0", "H2", "H3", "H4", "H5", "H6", "H7")) +
  scale_y_continuous(name = "SNR (dB)") +
  scale_color_manual(values = c("Conv" = "#E31A1C", "Rapid" = "#1F78B4")) +
  ggtitle("Final Model: Quadratic Fit of SNR across Harmonics") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  labs(caption = "Circles = Mean, X = Median, Points = Individual observations\nLines represent quadratic model predictions")

final_plot_partial
```

=============================================================================
SESSION INFORMATION
=============================================================================

```{r session_info}
# Print session information for reproducibility
sessionInfo()
```
