---
title: "Conv - Rapid Correlation Analysis"
author: "Jonas"
date: "2025-03-22"
output: html_document
---

Performs correlation and Intraclass Correlation Coefficient (ICC) analyses to assess the agreement between Rapid and Conventional FFR recording methods.
Research Question 2: How strongly are Rapid and Conventional FFR responses correlated?

=============================================================================
SETUP AND CONFIGURATION
=============================================================================

```{r setup, include=FALSE}
# Clear workspace for clean start
rm(list = ls())

# Load required libraries with error handling
required_packages <- c("tidyverse", "broom", "irr")

lapply(required_packages, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
})

# Set global constants
HARMONICS_OF_INTEREST <- c("F0", "H2", "H3", "H4", "H5", "H6", "H7")
HARMONIC_LABELS       <- c("F0 (EFR)", "H2 (EFR)", "H3 (EFR)", "H4 (TFS)", 
                           "H5 (TFS)", "H6 (TFS)", "H7 (TFS)")
TARGET_RUN            <- "Run1"
TARGET_ITERATION      <- 100
```

=============================================================================
DATA LOADING AND PREPROCESSING
=============================================================================

```{r data_loading}
# Define file paths (consider using here::here() for better path management)
DATA_PATH   <- "./input/"
SNR_FILE    <- "SNRIterationsMatrixClean.csv"
SIGNAL_FILE <- "SignalIterationsMatrixClean.csv"

# Load and preprocess SNR data
load_and_clean_data <- function(file_path) {
  data <- read_csv(file_path, show_col_types = FALSE) 
  return(data)
}

snr_data_raw    <- load_and_clean_data(file.path(DATA_PATH, SNR_FILE))
signal_data_raw <- load_and_clean_data(file.path(DATA_PATH, SIGNAL_FILE))

# Calculate time matching between Rapid and Conventional conditions
# This ensures fair comparison by matching recording durations
calculate_time_matching <- function(data) {
  
  # Get maximum recording time for Rapid condition
  rapid_max_time <- data %>%
    filter(listener  == 1, 
           condition == "Rapid", 
           run       == "Run1", 
           harmonic  == "F0") %>%
    pull(recording_time) %>%
    max(na.rm = TRUE)
  
  # Find closest matching time point in Conventional condition
  conv_matching_times <- data %>%
    filter(listener   == 1, 
           condition  == "Conv", 
           run        == "Run1", 
           harmonic   == "F0") %>%
    mutate(time_diff  = abs(recording_time - rapid_max_time)) %>%
    filter(time_diff  == min(time_diff, na.rm = TRUE))
  
  recording_time_match   <- conv_matching_times$recording_time[1]
  stimulation_time_match <- conv_matching_times$stimulation_time[1]
  
  return(list(
    rapid_max_time         = rapid_max_time,
    recording_time_match   = recording_time_match,
    stimulation_time_match = stimulation_time_match
  ))
}

time_matching <- calculate_time_matching(snr_data_raw)
```

=============================================================================
UTILITY FUNCTIONS
=============================================================================

```{r utility_functions}
# Function to convert harmonic names to numeric values
convert_harmonics_to_numeric <- function(data) {
  data %>%
    mutate(
      harmonic = recode_factor(
        harmonic,
        "F0" = 1, "H2" = 2, "H3" = 3, "H4" = 4,
        "H5" = 5, "H6" = 6, "H7" = 7
      ),
      harmonic = as.numeric(harmonic)
    )
}

# Function to prepare statistics data
prepare_stats_data <- function(data, signal_filter = NULL) {
  filtered_data <- data %>%
    filter(
      harmonic %in% HARMONICS_OF_INTEREST,
      run == TARGET_RUN,
      (condition == "Rapid" & iteration == TARGET_ITERATION) |
      (condition == "Conv" & iteration == TARGET_ITERATION)
    )
  
  # Apply signal filter if provided
  if (!is.null(signal_filter)) {
    filtered_data <- filtered_data %>%
      filter(signal_noise == signal_filter)
  }
  
  filtered_data %>%
    convert_harmonics_to_numeric() %>%
    group_by(condition, run, harmonic) %>%
    mutate(
      mean_value = mean(data_value, na.rm = TRUE),
      median_value = median(data_value, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(listener, run, condition, harmonic, data_value, mean_value, median_value)
}

# Function to create correlation plot
create_correlation_plot <- function(data, x_label, y_label, x_pos = -15, y_pos = 19) {
  # Calculate summary statistics
  summary_stats <- data %>%
    group_by(harmonic) %>%
    summarise(
      sd_conv = sd(Conv, na.rm = TRUE),
      sd_rapid = sd(Rapid, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      label = paste0("SD for Conv: ", round(sd_conv, 3), 
                    "\nSD for Rapid: ", round(sd_rapid, 3))
    )
  
  # Create plot
  ggplot(data, aes(x = Rapid, y = Conv)) +
    geom_point() +
    geom_smooth(method = "lm") +
    geom_abline(intercept = 0, slope = 1, color = "grey", 
                linetype = "dashed", size = 0.5) +
    scale_x_continuous(name = x_label) +
    scale_y_continuous(name = y_label) +
    geom_text(
      data = summary_stats,
      aes(x = x_pos, y = y_pos, label = label),
      hjust = 0, vjust = 0, inherit.aes = FALSE, size = 2
    ) +
    facet_wrap(~harmonic) +
    theme_bw()
}

# Function to calculate ICC for a single harmonic
calculate_harmonic_icc <- function(data, harmonic_num) {
  harmonic_data <- data %>%
    filter(harmonic == harmonic_num) %>%
    select(Rapid, Conv)
  
  if (nrow(harmonic_data) == 0) {
    warning(paste("No data found for harmonic", harmonic_num))
    return(NULL)
  }
  
  icc(harmonic_data, model = "oneway")
} #return??

# Function to extract ICC results in a structured format
extract_icc_results <- function(icc_result) {
  if (is.null(icc_result)) {
    return(list(value = NA, df1 = NA, df2 = NA, lbound = NA, ubound = NA, p.value = NA))
  }
  
  list(
    value = icc_result$value,
    df1 = icc_result$df1,
    df2 = icc_result$df2,
    lbound = icc_result$lbound,
    ubound = icc_result$ubound,
    p.value = icc_result$p.value
  )
}
```

=============================================================================
DATA PREPARATION
=============================================================================

```{r data_preparation}
# Prepare SNR statistics data
snr_stats_data <- prepare_stats_data(snr_data_raw)

# Prepare Signal statistics data  
signal_stats_data <- prepare_stats_data(signal_data_raw, "Signal")

cat("Data preparation completed:\n")
cat("SNR data dimensions:", dim(snr_stats_data), "\n")
cat("Signal data dimensions:", dim(signal_stats_data), "\n")
```

=============================================================================
CORRELATION AND VISUALISATIONS
=============================================================================

```{r correlation_plots}
# Prepare data for correlation plots
snr_correlation_data <- snr_stats_data %>%
  select(listener, condition, harmonic, data_value) %>%
  pivot_wider(names_from = condition, values_from = data_value)

signal_correlation_data <- signal_stats_data %>%
  select(listener, condition, harmonic, data_value) %>%
  pivot_wider(names_from = condition, values_from = data_value)

# Create SNR correlation plot
snr_correlation_plot <- create_correlation_plot(
  snr_correlation_data,
  "Rapid FFR SNR (dB)",
  "Conventional FFR SNR (dB)",
  x_pos = -15, y_pos = 19
)

# Create Signal correlation plot
signal_correlation_plot <- create_correlation_plot(
  signal_correlation_data,
  "Rapid FFR RMS amplitude (μV RMS)",
  "Conventional FFR RMS amplitude (μV RMS)",
  x_pos = 0.135, y_pos = 0.01
)

# Display plots
snr_correlation_plot
signal_correlation_plot
```

=============================================================================
ICC ANALYSIS FOR SIGNAL DATA
=============================================================================

```{r icc_analysis}
# Calculate ICC for all harmonics using lapply instead of for loop
icc_results <- lapply(1:7, function(i) {
  calculate_harmonic_icc(signal_correlation_data, i)
})

# Name the results for easier access
names(icc_results) <- paste0("ICCHarmonic", 1:7)

# Extract p-values for multiple comparison correction
p_values <- sapply(icc_results, function(x) {
  if (is.null(x)) NA else x$p.value
})

# Apply Holm correction for multiple comparisons
adjusted_p_values <- p.adjust(p_values, method = "holm", n = length(p_values))

# Display results
cat("Original p-values:\n")
round(p_values, 4)
cat("\nHolm-adjusted p-values:\n")
round(adjusted_p_values, 4)

# Extract structured results for all harmonics
icc_summary <- map_dfr(1:7, function(i) {
  result <- extract_icc_results(icc_results[[i]])
  tibble(
    Harmonic = i,
    Harmonic_Label = HARMONIC_LABELS[i],
    ICC_Value = result$value,
    DF1 = result$df1,
    DF2 = result$df2,
    CI_Lower = result$lbound,
    CI_Upper = result$ubound,
    P_Value = result$p.value,
    P_Value_Adjusted = adjusted_p_values[i]
  )
})

# Display summary table
icc_summary
```

=============================================================================
RESULTS TABLE
=============================================================================

```{r results_table}
# Generate formatted results table for LaTeX/publication
generate_latex_table <- function(icc_summary) {
  icc_summary %>%
    mutate(
      ICC_Formatted = round(ICC_Value, 3),
      DF_Formatted = paste0("[", round(DF1, 0), ", ", round(DF2, 0), "]"),
      CI_Formatted = paste0("[", round(CI_Lower, 3), ", ", round(CI_Upper, 3), "]"),
      P_Formatted = round(P_Value, 3),
      P_Adj_Formatted = round(P_Value_Adjusted, 3)
    ) %>%
    select(Harmonic_Label, ICC_Formatted, DF_Formatted, CI_Formatted, 
           P_Formatted, P_Adj_Formatted)
}

latex_table <- generate_latex_table(icc_summary)
latex_table

# Create publication-ready table
cat("\nPublication-ready LaTeX table rows:\n")
for (i in 1:nrow(latex_table)) {
  cat(paste0(
    latex_table$Harmonic_Label[i], " & ",
    latex_table$ICC_Formatted[i], " & ",
    latex_table$DF_Formatted[i], " & ",
    latex_table$CI_Formatted[i], " & ",
    latex_table$P_Formatted[i], " & ",
    latex_table$P_Adj_Formatted[i], " \\\\ \n"
  ))
}
```

=============================================================================
ADDITIONAL ANALYSIS
=============================================================================

```{r diagnostic_analysis}
# Create diagnostic scatter plots for each harmonic
diagnostic_plots <- map(1:7, function(i) {
  harmonic_data <- signal_correlation_data %>%
    filter(harmonic == i)
  
  if (nrow(harmonic_data) == 0) return(NULL)
  
  ggplot(harmonic_data, aes(x = Rapid, y = Conv)) +
    geom_point(alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE) +
    geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
    labs(
      title = paste("Harmonic", i, "-", HARMONIC_LABELS[i]),
      x = "Rapid FFR (μV RMS)",
      y = "Conventional FFR (μV RMS)"
    ) +
    theme_bw()
})

# Display individual diagnostic plots
walk(diagnostic_plots, print)

# Calculate additional correlation statistics
correlation_stats <- map_dfr(1:7, function(i) {
  harmonic_data <- signal_correlation_data %>%
    filter(harmonic == i) %>%
    select(Rapid, Conv) %>%
    drop_na()
  
  if (nrow(harmonic_data) < 3) {
    return(tibble(Harmonic = i, Correlation = NA, P_Value = NA, N = 0))
  }
  
  cor_test <- cor.test(harmonic_data$Rapid, harmonic_data$Conv)
  
  tibble(
    Harmonic = i,
    Harmonic_Label = HARMONIC_LABELS[i],
    Correlation = cor_test$estimate,
    P_Value = cor_test$p.value,
    CI_Lower = cor_test$conf.int[1],
    CI_Upper = cor_test$conf.int[2],
    N = nrow(harmonic_data)
  )
})

cat("\nPearson Correlation Results:\n")
correlation_stats
```

=============================================================================
SUMMARY
=============================================================================

```{r summary}
# Combine ICC and correlation results
combined_results <- icc_summary %>%
  left_join(correlation_stats, by = c("Harmonic", "Harmonic_Label")) %>%
  select(
    Harmonic_Label, N, ICC_Value, CI_Lower.x, CI_Upper.x, P_Value.x,
    Correlation, CI_Lower.y, CI_Upper.y, P_Value.y, P_Value_Adjusted
  ) %>%
  rename(
    ICC = ICC_Value,
    ICC_CI_Lower = CI_Lower.x,
    ICC_CI_Upper = CI_Upper.x,
    ICC_P_Value = P_Value.x,
    Correlation_CI_Lower = CI_Lower.y,
    Correlation_CI_Upper = CI_Upper.y,
    Correlation_P_Value = P_Value.y,
    ICC_P_Adjusted = P_Value_Adjusted
  )

cat("\nCombined Results Summary:\n")
combined_results

# Identify significant results
significant_icc <- combined_results %>%
  filter(ICC_P_Adjusted < 0.05) %>%
  pull(Harmonic_Label)

significant_correlation <- combined_results %>%
  filter(Correlation_P_Value < 0.05) %>%
  pull(Harmonic_Label)

cat("\nSignificant ICC results (adjusted p < 0.05):", 
    if(length(significant_icc) > 0) paste(significant_icc, collapse = ", ") else "None", "\n")
cat("Significant correlations (p < 0.05):", 
    if(length(significant_correlation) > 0) paste(significant_correlation, collapse = ", ") else "None", "\n")
```

=============================================================================
SESSION INFORMATION
=============================================================================

```{r session_info}
# Print session information for reproducibility
sessionInfo()
```