---
title: "Test-Retest Reliability Analysis: Conventional vs Rapid FFR"
author: "Jonas"
date: "2025-03-22"
output: html_document
---

Evaluates test-retest reliability of FFR measurements across Rapid and Conventional methods using ICC and correlation metrics.
Research Question 3: Are there differences in test-retest reliability (ICC) between the Rapid and Conventional FFR approaches?

=============================================================================
SETUP AND CONFIGURATION
=============================================================================

```{r setup, include=FALSE}
# Clear workspace for clean start
rm(list = ls())

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load_packages, include=FALSE}
# Load required libraries with error handling
required_packages <- c("tidyverse", "irr")

lapply(required_packages, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
})
```

```{r load_data, include=FALSE}
# File paths - consider using here::here() for better path management
data_path <-  "./input/"

# Load SNR data
snr_data <- read_csv(file.path(data_path, "SNRIterationsMatrixClean.csv"))

# Load Signal data  
signal_data <- read_csv(file.path(data_path, "SignalIterationsMatrixClean.csv"))

```

=============================================================================
DATA PREPROCESSING
=============================================================================

```{r data_preprocessing}
# Constants for harmonics of interest
HARMONICS_OF_INTEREST   <- c('F0', 'H2', 'H3', 'H4', 'H5', 'H6', 'H7')
TARGET_ITERATION_RAPID  <- 100

# Find matching recording time between Rapid and Conventional conditions
find_matching_recording_time <- function(snr_data) {
  # Get maximum recording time for Rapid condition
  rapid_max_time <- snr_data %>%
    filter(
      listener == 1,
      condition == "Rapid",
      run == "Run1",
      harmonic == "F0"
    ) %>%
    pull(recording_time) %>%
    max()
  
  # Find closest matching time in Conventional condition
  matching_time <- snr_data %>%
    filter(
      listener == 1,
      condition == "Conv",
      run == "Run1",
      harmonic == "F0"
    ) %>%
    mutate(time_diff = abs(recording_time - rapid_max_time)) %>%
    filter(time_diff == min(time_diff)) %>%
    pull(recording_time) %>%
    first()
  
  return(matching_time)
}

recording_time_match <- find_matching_recording_time(snr_data)
```

```{r prepare_analysis_data}
# Prepare conventional FFR data for analysis
prepare_conventional_data <- function(signal_data, recording_time_match) {
  signal_data %>%
    filter(
      signal_noise == "Signal",
      harmonic %in% HARMONICS_OF_INTEREST,
      recording_time == recording_time_match,
      condition == "Conv"
    ) %>%
    group_by(condition, run, harmonic) %>%
    mutate(
      mean_value = mean(data_value, na.rm = TRUE),
      median_value = median(data_value, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(listener, run, condition, harmonic, data_value, mean_value, median_value)
}

# Prepare rapid FFR data for analysis
prepare_rapid_data <- function(signal_data) {
  signal_data %>%
    filter(
      signal_noise == "Signal",
      harmonic %in% HARMONICS_OF_INTEREST,
      iteration == TARGET_ITERATION_RAPID,
      condition == "Rapid"
    ) %>%
    group_by(condition, run, harmonic) %>%
    mutate(
      mean_value = mean(data_value, na.rm = TRUE),
      median_value = median(data_value, na.rm = TRUE)
    ) %>%
    ungroup() %>%
    select(listener, run, condition, harmonic, data_value, mean_value, median_value)
}

rapid_data        <- prepare_rapid_data(signal_data)
conventional_data <- prepare_conventional_data(signal_data, recording_time_match)
```

=============================================================================
VISUALISATION
=============================================================================

```{r visualization_functions}
# Create test-retest scatter plots with summary statistics
create_test_retest_plot <- function(data, condition_name) {
  # Prepare data for plotting
  plot_data <- data %>%
    select(listener, condition, run, harmonic, data_value) %>%
    pivot_wider(names_from = "run", values_from = "data_value")
  
  # Calculate summary statistics
  summary_stats <- plot_data %>%
    group_by(harmonic) %>%
    summarise(
      sd_run1 = sd(Run1, na.rm = TRUE),
      sd_run2 = sd(Run2, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    mutate(
      label = paste0(
        "SD Run 1: ", round(sd_run1, 3), 
        "\nSD Run 2: ", round(sd_run2, 3)
      )
    )
  
  # Create plot
  ggplot(plot_data, aes(x = Run1, y = Run2)) + 
    geom_point(alpha = 0.7) +
    geom_smooth(method = 'lm', se = TRUE, color = "blue") +
    geom_abline(intercept = 0, slope = 1, color = "grey", linetype = "dashed", size = 0.5) +
    geom_text(
      data = summary_stats,
      aes(x = 0.135, y = 0.01, label = label),
      hjust = 0, vjust = 0, inherit.aes = FALSE, size = 2
    ) +
    facet_wrap(~harmonic, labeller = label_both) +
    labs(
      x = paste("Run 1", condition_name, "FFR RMS amplitude (μV RMS)"),
      y = paste("Run 2", condition_name, "FFR RMS amplitude (μV RMS)"),
      title = paste("Test-Retest Reliability:", condition_name)
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(hjust = 0.5),
      strip.background = element_rect(fill = "lightgray")
    )
}
```

```{r create_plots}
# Generate test-retest plots
conventional_plot <- create_test_retest_plot(conventional_data, "Conventional")
rapid_plot        <- create_test_retest_plot(rapid_data, "Rapid")

# Display plots
conventional_plot
rapid_plot
```

=============================================================================
ICC ANALYSIS
=============================================================================

```{r icc_functions}
# Calculate ICC for a single condition and harmonic
calculate_single_icc <- function(data, harmonic_num) {
  icc_data <- data %>%
    filter(harmonic == harmonic_num) %>%
    select(Run1, Run2)
  
  if (nrow(icc_data) == 0) {
    return(NULL)
  }
  
  icc(icc_data, model = "twoway", type = "agreement")
}

# Calculate ICC for all harmonics in a condition
calculate_condition_icc <- function(data, condition_name) {
  # Prepare data in wide format
  icc_data <- data %>%
    select(listener, run, harmonic, data_value) %>%
    pivot_wider(names_from = "run", values_from = "data_value") %>%
    mutate(
      harmonic = recode_factor(
        harmonic,
        "F0" = 1, "H2" = 2, "H3" = 3, "H4" = 4,
        "H5" = 5, "H6" = 6, "H7" = 7
      ),
      harmonic = as.numeric(harmonic)
    )
  
  # Calculate ICC for each harmonic
  icc_results        <- map(1:7, ~ calculate_single_icc(icc_data, .x))
  names(icc_results) <- paste0(condition_name, "_ICC_Harmonic_", 1:7)
  
  # Extract p-values for multiple comparison correction
  p_values_raw       <- map_dbl(icc_results, ~ ifelse(is.null(.x), NA, .x$p.value))
  p_values_adjusted  <- p.adjust(p_values_raw, method = 'holm', n = length(p_values_raw))
  
  return(list(
    icc_results = icc_results,
    p_values_raw = p_values_raw,
    p_values_adjusted = p_values_adjusted
  ))
}
```

```{r run_icc_analysis}
# Calculate ICC for both conditions
rapid_icc_analysis        <- calculate_condition_icc(rapid_data, "Rapid")
conventional_icc_analysis <- calculate_condition_icc(conventional_data, "Conv")

# Display results
cat("Rapid FFR ICC p-values (raw):\n")
print(round(rapid_icc_analysis$p_values_raw, 4))
cat("\nRapid FFR ICC p-values (Holm-adjusted):\n")
print(round(rapid_icc_analysis$p_values_adjusted, 4))

cat("\n\nConventional FFR ICC p-values (raw):\n")
print(round(conventional_icc_analysis$p_values_raw, 4))
cat("\nConventional FFR ICC p-values (Holm-adjusted):\n")
print(round(conventional_icc_analysis$p_values_adjusted, 4))
```

=============================================================================
PERMUTATUION TESTING
=============================================================================

```{r permutation_functions}
# Perform permutation test comparing ICC values between conditions
perform_permutation_test <- function(rapid_data, conventional_data, n_permutations = 1000) {
  # Prepare data for permutation
  rapid_wide <- rapid_data %>%
    select(listener, run, harmonic, data_value) %>%
    pivot_wider(names_from = "run", values_from = "data_value") %>%
    mutate(condition = "Rapid") %>%
    mutate(
      harmonic = recode_factor(
        harmonic,
        "F0" = 1, "H2" = 2, "H3" = 3, "H4" = 4,
        "H5" = 5, "H6" = 6, "H7" = 7
      ),
      harmonic = as.numeric(harmonic)
    )
  
  conv_wide <- conventional_data %>%
    select(listener, run, harmonic, data_value) %>%
    pivot_wider(names_from = "run", values_from = "data_value") %>%
    mutate(condition = "Conv") %>%
    mutate(
      harmonic = recode_factor(
        harmonic,
        "F0" = 1, "H2" = 2, "H3" = 3, "H4" = 4,
        "H5" = 5, "H6" = 6, "H7" = 7
      ),
      harmonic = as.numeric(harmonic)
    )
  
  # Initialise results matrices
  n_harmonics              <- 7
  bootstrapped_differences <- matrix(nrow = n_permutations, ncol = n_harmonics)
  real_differences         <- numeric(n_harmonics)
  
  # Function to calculate ICC difference for a single harmonic
  calculate_icc_difference <- function(harmonic_num, data_rapid, data_conv) {
    rapid_icc <- calculate_single_icc(data_rapid, harmonic_num)
    conv_icc  <- calculate_single_icc(data_conv, harmonic_num)
    
    if (is.null(rapid_icc) || is.null(conv_icc)) {
      return(NA)
    }
    
    return(rapid_icc$value - conv_icc$value)
  }
  
  # Calculate real ICC differences
  for (harmonic_It in 1:n_harmonics) {
    real_differences[harmonic_It] <- calculate_icc_difference(harmonic_It, 
                                                           rapid_wide, conv_wide
    )
    
    # Perform permutation testing
    # Get harmonic-specific data
    harmonic_data_rapid <- rapid_wide %>% 
      filter(harmonic == harmonic_It) %>%
      select(Run1, Run2)
      
    harmonic_data_conv <- conv_wide %>% 
      filter(harmonic == harmonic_It) %>%
      select(Run1, Run2)
    
    for (perm_It in 1:n_permutations) {
      set.seed(perm_It)
      BootstrappedRapidRun1 <- sample (harmonic_data_rapid$Run1, 16, replace = FALSE)
    
      set.seed(perm_It+1001)
      BootstrappedRapidRun2 <- sample (harmonic_data_rapid$Run2, 16, replace = FALSE)
      
      # Get the conv-boostrap, while matching listeners with the rapid by using the same seed
      set.seed(perm_It)
      BootstrappedConvRun1 <- sample (harmonic_data_conv$Run1, 16, replace = FALSE)
    
      set.seed(perm_It+1001)
      BootstrappedConvRun2 <- sample (harmonic_data_conv$Run2, 16, replace = FALSE)
      
      # Compute the 2 ICCs, and then get the difference
      BoostrappedRapidJoined <- cbind (BootstrappedRapidRun1, BootstrappedRapidRun2)
      rapid_icc              <- icc(BoostrappedRapidJoined, model = "twoway", type = "agreement")
      
      BoostrappedConvJoined  <- cbind (BootstrappedConvRun1, BootstrappedConvRun2)
      conv_icc               <- icc(BoostrappedConvJoined, model = "twoway", type = "agreement")
      
      bootstrapped_differences[perm_It, harmonic_It] <- rapid_icc$value - conv_icc$value
    }
  }
  
  # Calculate p-values
  p_values_two_tailed <- sapply(1:n_harmonics, function(harmonic_It) {
    sum(abs(real_differences[harmonic_It]) < 
          abs(bootstrapped_differences[, harmonic_It])) / n_permutations
  })
  
  # Calculate confidence intervals
  confidence_intervals <- sapply(1:n_harmonics, function(harmonic_It) {
    quantile(bootstrapped_differences[, harmonic_It], c(0.025, 0.975))
    })
  
  return(list(
    real_differences = real_differences,
    bootstrapped_differences = bootstrapped_differences,
    p_values_two_tailed = p_values_two_tailed,
    confidence_intervals = confidence_intervals
  ))
}
```

```{r run_permutation_test}
# Perform permutation test
permutation_results <- perform_permutation_test(rapid_data, conventional_data, 1000)

cat("Permutation test results (two-tailed p-values):\n")
print(round(permutation_results$p_values_two_tailed, 4))

cat("\nReal ICC differences (Rapid - Conventional):\n")
print(round(permutation_results$real_differences, 4))
```

=============================================================================
PERMUTATUION TESTING VISUALISATION
=============================================================================

```{r visualisation_permutation}
# Create permutation test histograms
create_permutation_plots <- function(permutation_results) {
  n_harmonics <- ncol(permutation_results$bootstrapped_differences)
  
  plot_list <- map(1:n_harmonics, function(h) {
    # Prepare data for plotting
    boot_data <- data.frame(
      differences = permutation_results$bootstrapped_differences[, h]
    )
    
    # Get confidence intervals and real difference
    ci_lower   <- permutation_results$confidence_intervals[1, h]
    ci_upper   <- permutation_results$confidence_intervals[2, h]
    real_diff  <- permutation_results$real_differences[h]
    p_val      <- permutation_results$p_values_two_tailed[h]
    
    # Create plot
    ggplot(boot_data, aes(x = differences)) +
      geom_density(fill = "lightblue", alpha = 0.7) +
      geom_vline(xintercept = ci_lower, linetype = "dotted", color = "black", size = 0.5) +
      geom_vline(xintercept = ci_upper, linetype = "dotted", color = "black", size = 0.5) +
      geom_vline(xintercept = real_diff, linetype = "solid", color = "red", size = 1) +
      labs(
        x = "Rapid ICC - Conventional ICC",
        y = "Density",
        title = paste("Harmonic", h, "- p =", round(p_val, 4)),
        subtitle = paste("Real difference:", round(real_diff, 4))
      ) +
      theme_bw() +
      theme(
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)
      )
  })
  
  return(plot_list)
}

# Generate and display permutation plots
permutation_plots <- create_permutation_plots(permutation_results)

# Display plots
walk(permutation_plots, print)
```

=============================================================================
SUMMARY
=============================================================================

```{r summary_tables}
# Create publication-ready ICC summary table
create_icc_summary_table <- function(icc_analysis, condition_name) {
  icc_values <- map_dbl(icc_analysis$icc_results, ~ ifelse(is.null(.x), NA, .x$value))
  ci_lower   <- map_dbl(icc_analysis$icc_results, ~ ifelse(is.null(.x), NA, .x$lbound))
  ci_upper   <- map_dbl(icc_analysis$icc_results, ~ ifelse(is.null(.x), NA, .x$ubound))
  df1        <- map_dbl(icc_analysis$icc_results, ~ ifelse(is.null(.x), NA, .x$df1))
  df2        <- map_dbl(icc_analysis$icc_results, ~ ifelse(is.null(.x), NA, .x$df2))
  
  harmonic_labels <- c("F0 (EFR)", "H2 (EFR)", "H3 (EFR)", "H4 (TFS)", 
                      "H5 (TFS)", "H6 (TFS)", "H7 (TFS)")
  
  data.frame(
    Condition        = condition_name,
    Harmonic         = harmonic_labels,
    ICC              = round(icc_values, 3),
    CI_Lower         = round(ci_lower, 3),
    CI_Upper         = round(ci_upper, 3),
    DF1              = round(df1, 0),
    DF2              = round(df2, 0),
    p_value_raw      = round(icc_analysis$p_values_raw, 4),
    p_value_adjusted = round(icc_analysis$p_values_adjusted, 4)
  )
}

# Create summary tables
rapid_summary <- create_icc_summary_table(rapid_icc_analysis, "Rapid")
conv_summary <- create_icc_summary_table(conventional_icc_analysis, "Conventional")

# Display tables
cat("Rapid FFR ICC Summary:\n")
print(rapid_summary)

cat("\n\nConventional FFR ICC Summary:\n")
print(conv_summary)
```

=============================================================================
SESSION INFORMATION
=============================================================================

```{r session_info}
# Document the R session for reproducibility
sessionInfo()
```