---
title: "FFR Experiment 1 Plots"
author: "Jonas"
date: "2025-03-25"
output: html_document
---

Generates all visualisations for Experiment 1, comparing FFR data obtained via Rapid and Conventional recording methods across multiple harmonics.

============================================================================
SETUP AND DEPENDENCIES
============================================================================

```{r setup, include=FALSE}
# Clear workspace for clean start
rm(list = ls())

# Load required libraries with error handling
required_packages <- c("tidyverse", "ggpubr")

lapply(required_packages, function(pkg) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg)
    library(pkg, character.only = TRUE)
  }
})

# Set global constants
NUM_PARTICIPANTS      <- 16
ALPHA_LEVEL           <- 0.05
HARMONICS_OF_INTEREST <- c("F0", "H2", "H3", "H4", "H5", "H6", "H7")
```

============================================================================
DATA LOADING & PROCESSING
============================================================================

```{r data_loading}
# Define file paths (consider moving to config file)
data_path   <- "./input/"
output_path <- "./output/"

# Function to load and preprocess data
load_and_preprocess_data <- function(filename) {
  file_path <- file.path(data_path, filename)
  
  read_csv(file_path, show_col_types = FALSE) 
}

# Load both datasets
snr_data    <- load_and_preprocess_data("SNRIterationsMatrixClean.csv")
signal_data <- load_and_preprocess_data("SignalIterationsMatrixClean.csv")

# Calculate time matching parameters for Rapid vs Conventional comparison
rapid_max_time <- snr_data %>%
  filter(listener == 1, condition == "Rapid", run == "Run1", harmonic == "F0") %>%
  pull(recording_time) %>%
  max()

# Find matching recording time in Conventional condition
matched_recording_time <- snr_data %>%
  filter(listener == 1, condition == "Conv", run == "Run1", harmonic == "F0") %>%
  mutate(time_diff = abs(recording_time - rapid_max_time)) %>%
  filter(time_diff == min(time_diff)) %>%
  pull(recording_time) %>%
  first()

# Find matching stimulation time
matched_stimulation_time <- snr_data %>%
  filter(listener == 1, condition == "Conv", run == "Run1", harmonic == "F0") %>%
  mutate(time_diff = abs(stimulation_time - rapid_max_time)) %>%
  filter(time_diff == min(time_diff)) %>%
  pull(stimulation_time) %>%
  first()

cat("Rapid max time:", rapid_max_time, "\n")
cat("Matched recording time:", matched_recording_time, "\n")
cat("Matched stimulation time:", matched_stimulation_time, "\n")
```

============================================================================
UTILITY FUNCTION
============================================================================

```{r utility_functions}
# Function to calculate confidence intervals
calculate_confidence_intervals <- function(data, value_col, n = NUM_PARTICIPANTS) {
  data %>%
    summarise(
      mean_value = mean({{value_col}}, na.rm = TRUE),
      sd_value   = sd({{value_col}}, na.rm = TRUE),
      ci_lower   = mean_value - qt(0.975, df = n - 1) * sd_value / sqrt(n),
      ci_upper   = mean_value + qt(0.975, df = n - 1) * sd_value / sqrt(n),
      .groups    = "keep"
    )
}

# Function to add time columns based on condition
add_time_columns <- function(data) {
  data %>%
    mutate(
      recording_time_mins = case_when(
        str_starts(condition, "C") ~ iteration * (5/100),
        str_starts(condition, "R") ~ iteration * (1.933333/100),
        TRUE ~ NA_real_
      ),
      stimulation_time_mins = case_when(
        str_starts(condition, "C") ~ iteration * (2.734375/100),
        str_starts(condition, "R") ~ iteration * (1.933333/100),
        TRUE ~ NA_real_
      )
    )
}

# Function to create growth plots
create_growth_plot <- function(data, x_var, x_label, title_suffix = "") {
  ggplot(data, aes(x = {{x_var}}, y = mean_value, 
                   group = condition, color = condition)) +
    geom_smooth(aes(ymin = ci_lower, ymax = ci_upper, fill = condition), 
                stat = "identity") +
    scale_x_continuous(name = x_label) +
    scale_y_continuous(name = "SNR (dB)") +
    theme_bw() +
    theme(legend.position = "bottom")
}

# Function to save plots with consistent parameters
save_plot_both_formats <- function(plot_obj, filename, width = 20, height = 10) {
  # Save as PNG
  ggsave(
    plot = plot_obj,
    filename = file.path(output_path, paste0(filename, ".png")),
    device = "png", dpi = 300, width = width, height = height, units = "cm"
  )
  
  # Save as TIFF
  ggsave(
    plot = plot_obj,
    filename = file.path(output_path, paste0(filename, ".tiff")),
    device = "tiff", dpi = 1200, width = width, height = height, units = "cm"
  )
}
```

============================================================================
F0 ANALYSIS: RECORDING/STIMULATIOIN TIME GROWTH
============================================================================

```{r f0_analysis}
# Prepare F0 growth data
f0_growth_data <- snr_data %>%
  filter(harmonic == "F0", run == "Run1") %>%
  group_by(condition, iteration) %>%
  calculate_confidence_intervals(data_value) %>%
  add_time_columns()

# Create F0 recording time growth plot
f0_recording_plot <- create_growth_plot(
  f0_growth_data, 
  recording_time_mins, 
  "Elapsed Time (mins)"
)

# Create F0 stimulation time growth plot  
f0_stimulation_plot <- create_growth_plot(
  f0_growth_data, 
  stimulation_time_mins, 
  "Stimulation Time (mins)"
)

# Combine plots
f0_combined_plot <- ggarrange(
  f0_recording_plot, f0_stimulation_plot,
  labels = c("A", "B"),
  common.legend = TRUE,
  legend = "bottom",
  ncol = 2
)

# Display plots
f0_recording_plot
f0_stimulation_plot
f0_combined_plot
```

============================================================================
ALL HARMONICS ANALYSIS: RECORDING/STIMULATIOIN TIME GROWTH
============================================================================

```{r harmonics_analysis}
# Prepare harmonics growth data
harmonics_growth_data <- snr_data %>%
  filter(harmonic %in% HARMONICS_OF_INTEREST, run == "Run1") %>%
  group_by(condition, iteration, harmonic) %>%
  calculate_confidence_intervals(data_value) %>%
  add_time_columns()

# Create harmonics recording time plot
harmonics_recording_plot <- harmonics_growth_data %>%
  ggplot(aes(x = recording_time_mins, y = mean_value, 
             group = condition, color = condition)) +
  geom_smooth(aes(ymin = ci_lower, ymax = ci_upper, fill = condition), 
              stat = "identity") +
  scale_x_continuous(name = "Elapsed Time (mins)") +
  scale_y_continuous(name = "SNR (dB)") +
  facet_wrap(~harmonic) +
  theme_bw()

# Create harmonics stimulation time plot
harmonics_stimulation_plot <- harmonics_growth_data %>%
  ggplot(aes(x = stimulation_time_mins, y = mean_value, 
             group = condition, color = condition)) +
  geom_smooth(aes(ymin = ci_lower, ymax = ci_upper, fill = condition), 
              stat = "identity") +
  scale_x_continuous(name = "Stimulation Time (mins)") +
  scale_y_continuous(name = "SNR (dB)") +
  facet_wrap(~harmonic) +
  theme_bw()

harmonics_recording_plot
harmonics_stimulation_plot
```

============================================================================
BOXPLOT ANALYSIS: MATCHED RECORDING TIME
============================================================================

```{r boxplot_analysis}
# Prepare matched data for boxplot
matched_data <- snr_data %>%
  filter(
    harmonic %in% HARMONICS_OF_INTEREST,
    run == "Run1",
    (condition == "Rapid" & iteration == 100) |
    (condition == "Conv" & recording_time == matched_recording_time)
  )

# Create boxplot
boxplot_matched <- ggplot(matched_data, 
                         aes(x = harmonic, y = data_value, fill = condition)) +
  geom_boxplot(width = 0.5, outlier.shape = NA) +
  geom_jitter(color = "black", size = 0.7, alpha = 0.3, 
              position = position_jitterdodge(jitter.width = 0.01)) +
  scale_x_discrete(name = "Harmonic") +
  scale_y_continuous(name = "SNR (dB)") +
  theme_bw()

# Generate summary statistics
summary_stats <- matched_data %>%
  group_by(harmonic, condition) %>%
  summarise(
    mean_snr = mean(data_value, na.rm = TRUE),
    sd_snr = sd(data_value, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

boxplot_matched
summary_stats
```

============================================================================
CORRELATION ANALYSIS: SIGNAL RAPID VS CONVENTIONAL
============================================================================

```{r correlation_analysis}
# Prepare correlation data
correlation_data <- signal_data %>%
  filter(
    run          == "Run1",
    iteration    == 100,
    signal_noise == "Signal",
    harmonic %in% HARMONICS_OF_INTEREST
  ) %>%
  select(listener, condition, harmonic, data_value) %>%
  pivot_wider(names_from = condition, values_from = data_value)

# Calculate summary statistics for plot
signal_summary_stats <- correlation_data %>%
  group_by(harmonic) %>%
  summarise(
    sd_conv = sd(Conv, na.rm = TRUE),
    sd_rapid = sd(Rapid, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(label = paste0(round(sd_conv, 3), "/", round(sd_rapid, 3)))

# Create basic correlation plot
correlation_plot_basic <- correlation_data %>%
  ggplot(aes(x = Rapid, y = Conv)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, color = "grey", 
              linetype = "dashed", size = 0.5) +
  scale_x_continuous(name = "Rapid FFR Signal Amplitude (μV RMS)") +
  scale_y_continuous(name = "Conventional FFR Signal Amplitude (μV RMS)") +
  facet_wrap(~harmonic) +
  theme_bw()

# Create enhanced quadratic correlation plot
correlation_plot_enhanced <- correlation_data %>%
  ggplot(aes(x = Rapid, y = Conv)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, color = "grey", 
              linetype = "dashed", size = 0.5) +
  scale_x_continuous(name = "Rapid FFR Signal Amplitude (μV RMS)", 
                     limits = c(0, 0.22)) +
  scale_y_continuous(name = "Conventional FFR Signal Amplitude (μV RMS)", 
                     limits = c(0, 0.22)) +
  geom_text(data = tibble(harmonic = HARMONICS_OF_INTEREST),
            aes(x = 0.01, y = 0.2, label = harmonic),
            hjust = 0, vjust = 0, inherit.aes = FALSE, size = 3) +
  geom_text(data = signal_summary_stats,
            aes(x = 0.135, y = 0.01, label = label),
            hjust = 0, vjust = 0, inherit.aes = FALSE, size = 2) +
  facet_wrap(~harmonic) +
  theme_bw() +
  theme(
    aspect.ratio = 1,
    axis.text.x = element_text(size = 7),
    strip.text.x = element_blank()
  )

correlation_plot_basic
correlation_plot_enhanced
```

============================================================================
TEST-RETEST RELIABILITY ANALYSIS
============================================================================

```{r test_retest_analysis}
# Function to create test-retest data
create_test_retest_data <- function(data, condition_filter, time_filter = NULL) {
  filtered_data <- data %>%
    filter(
      harmonic %in% HARMONICS_OF_INTEREST,
      condition == condition_filter,
      signal_noise == "Signal"
    )
  
  if (!is.null(time_filter)) {
    filtered_data <- filtered_data %>%
      filter(recording_time == time_filter)
  }else{
    filtered_data <- filtered_data %>%
      filter(iteration == 100)
  }
  
  filtered_data %>%
    select(listener, run, harmonic, data_value) %>%
    pivot_wider(names_from = run, values_from = data_value)
}

# Function to create test-retest plot
create_test_retest_plot <- function(data, x_label, y_label) {
  data %>%
    ggplot(aes(x = Run1, y = Run2)) +
    geom_point() +
    geom_smooth(method = "lm") +
    geom_abline(intercept = 0, slope = 1, color = "grey", 
                linetype = "dashed", size = 0.5) +
    scale_x_continuous(name = x_label) +
    scale_y_continuous(name = y_label) +
    geom_text(data = tibble(harmonic = HARMONICS_OF_INTEREST),
              aes(x = 0.01, y = 0.2, label = harmonic),
              hjust = 0, vjust = 0, inherit.aes = FALSE, size = 3) +
    facet_wrap(~harmonic) +
    theme_bw() +
    theme(
      aspect.ratio = 1,
      axis.text.x = element_text(size = 7),
      strip.text.x = element_blank()
    )
}

# Create Rapid test-retest data and plot
rapid_test_retest_data <- create_test_retest_data(signal_data, "Rapid")
rapid_test_retest_plot <- create_test_retest_plot(
  rapid_test_retest_data,
  "Run 1 Rapid FFR Signal Amplitude (μV RMS)",
  "Run 2 Rapid FFR Signal Amplitude (μV RMS)"
)

# Create Conventional test-retest data and plot (time-matched)
conv_test_retest_data <- create_test_retest_data(
  signal_data, "Conv", matched_recording_time
)
conv_test_retest_plot <- create_test_retest_plot(
  conv_test_retest_data,
  "Run 1 Conv FFR Signal Amplitude (μV RMS)",
  "Run 2 Conv FFR Signal Amplitude (μV RMS)"
)

# Create combined test-retest plot
x_axis_ticks <- c(0.0, 0.05, 0.1, 0.15, 0.2)
combined_test_retest_plot <- ggarrange(
  rapid_test_retest_plot + 
    scale_x_continuous(labels = x_axis_ticks,
                       name = "Run 1 Rapid FFR Signal Amplitude (μV RMS)"),
  conv_test_retest_plot + 
    scale_x_continuous(labels = x_axis_ticks,
                       name = "Run 1 Conv FFR Signal Amplitude (μV RMS)"),
  labels = c("A", "B"),
  common.legend = TRUE,
  legend = "bottom",
  ncol = 2
)

rapid_test_retest_plot
conv_test_retest_plot
combined_test_retest_plot
```

============================================================================
SAVE ALL PLOTS
============================================================================
```{r save_plots}
# List of plots to save with their corresponding names and dimensions
plots_to_save <- list(
  list(plot = f0_recording_plot, name = "F0RecordingTimeGrowthPlot", 
       width = 20, height = 10),
  list(plot = f0_stimulation_plot, name = "F0StimulationTimeGrowthPlot", 
       width = 20, height = 10),
  list(plot = f0_combined_plot, name = "F0GrowthPlot", 
       width = 20, height = 10),
  list(plot = harmonics_recording_plot, name = "HarmonicsSNRRecordingTimeGrowthPlotMean", 
       width = 20, height = 10),
  list(plot = harmonics_stimulation_plot, name = "HarmonicsSNRStimulationTimeGrowthPlotMean", 
       width = 20, height = 10),
  list(plot = boxplot_matched, name = "BoxplotMatchedRecordingTime", 
       width = 20, height = 10),
  list(plot = correlation_plot_enhanced, name = "CorrSignalRapidConvQuadratic", 
       width = 12, height = 12),
  list(plot = rapid_test_retest_plot, name = "RapidSignalRun1Run2", 
       width = 12, height = 12),
  list(plot = conv_test_retest_plot, name = "ConvSignalRun1Run2Matched", 
       width = 12, height = 12),
  list(plot = combined_test_retest_plot, name = "TestRetestPlot", 
       width = 24, height = 12)
)

# Save all plots using lapply instead of manual repetition
lapply(plots_to_save, function(plot_info) {
  save_plot_both_formats(
    plot_info$plot, 
    plot_info$name, 
    plot_info$width, 
    plot_info$height
  )
})

cat("All plots saved successfully to:", output_path, "\n")
```

============================================================================
SESSION INFORMATION
============================================================================
```{r session_info}
# Print session information for reproducibility
sessionInfo()
```